### 2016-05-26 thesis_paper/source.tex
adding some stuff to thesis paper

\documentclass[11pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage[dvips]{graphicx}
\usepackage{xcolor}
\usepackage{times}
\usepackage{ragged2e}
\usepackage{titlesec}
\usepackage{wrapfig,subfigure,setspace,verbatim,cite,url,graphicx,float}
\usepackage{algorithmic,amsfonts,amssymb,amsmath,calc,calrsfs,cleveref}
\usepackage[justification=centering]{caption}
\usepackage{url}

@@

Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams. Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices. 
\subsection{One-body kernel} Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages. Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (e.g. generic CUDA function atomicAdd). We run the first stream with one-body queries and wait till it gets finished. The reason is that we expect certain behavior from shared memory from each of the streams, thus, just to make sure that they don't overlap, we run them separately. While the computation is running, there is a possibility to use a third stream to load another chunk of data, too. First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep it in the GPU.
\subsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream. As it has already been mentioned, the biggest workload is in this kernel, because it has higher complexity. In this particular case it is $O(n^2)$. GPU streaming multiprocessors are very powerful for straightforward primitive independent calculations, and each core is much weaker than a moden CPU which has a whole bunch of L caches with some prediction behavior systems for code divergence. As you can see on Figure \ref{fg:threadload}, unbalanced thread load assumes for GPU kernel to check every time if the target atom can interact with other atom and if it has to wait till other threads finish their jobs. Thus, we had to come up with a strategy of making equally balanced thread jobs and distribute them among the threads leaving atom to thread mapping as is.

>>>>

\documentclass[11pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage[dvips]{graphicx}
\usepackage{xcolor}
\usepackage{times}
\usepackage{ragged2e}
\usepackage{titlesec}
\usepackage{wrapfig,subfigure,setspace,verbatim,cite,url,graphicx,float}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{amsfonts,amssymb,calc,calrsfs,cleveref}
\usepackage[justification=centering]{caption}
\usepackage{url}

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

@@

Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams. Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices. 
\subsection{One-body kernel}

As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach. Our implementation of one-body functions are relatively easy to understand.

\subsubsection{One-body functions implementation in CPU version}

Basically being given a vector of atoms in CPU version implementation (the one that we use for speedup benchmark) we loop through every atom and accumulate all the needed information to RAM. Once we have the result for one-body functions, we take it out for further usage, it can as well be used for further calculations where there are dependencies on such things as total mass, total charge, etc.

\vspace{1cm}

\begin{algorithm}
\caption{One body functions implementation in CPU version}
\label{alg:onebodycpu}
\begin{algorithmic}[1]
\Procedure{OneBodyFuncsCalc}{atoms, atomsCount}
\State \Comment{ $atoms$ is vector of atoms}
\State \Comment{ $atomsCount$ is number of atoms in the frame}

\State $\textit{results} \gets \textit{default initial results value}$

\BState \emph{top}:
\State $\textit{i} \gets \textit{0}$

\While{$i\not=atomsCount$} \Comment{Loop through all the atoms}

\State{$\textit{currentAtom} \gets \textit{atoms[i]}$}
\State{$\textit{calcTotalMass(results, currentAtom)}$}
\State{$\textit{calcTotalCharge(results, currentAtom)}$}
\State{$\textit{calcInertiaX(results, currentAtom)}$}
\State{$\textit{calcInertiaY(results, currentAtom)}$}
\State{$\textit{calcInertiaZ(results, currentAtom)}$}
\State{$\textit{calcDepoleMoment(results, currentAtom)}$}

\State{$\textit{//Here we are adding any other one-body functions}$}

%increment

\State $\textit{i} \gets \textit{i + 1}$ \Comment{Increment i}

\EndWhile

\State \textbf{return} $results$ \Comment{Return result for the frame}
\EndProcedure

\Procedure{calcTotalMass}{results, currentAtom} \Comment{Implementation of total mass function}
	\State $results \gets \textit{results.totalmass + currentAtom.mass}$ 
\EndProcedure

\Comment{All other functions}

\end{algorithmic}
\end{algorithm}

\vspace{1cm}

On Algorithm Pseudocode~\ref{alg:onebodycpu} you can see more detailed explanation of CPU version approach for calculating one-body functions.

\subsubsection{One-body functions implementation in GPU version}

Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages. Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (e.g. generic CUDA function atomicAdd). We run the first stream with one-body queries and wait till it gets finished. The reason is that we expect certain behavior from shared memory from each of the streams, thus, just to make sure that they don't overlap, we run them separately. While the computation is running, there is a possibility to use a third stream to load another chunk of data, too. First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep it in the GPU.

\subsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream. As it has already been mentioned, the biggest workload is in this kernel, because it has higher complexity. In this particular case it is $O(n^2)$. GPU streaming multiprocessors are very powerful for straightforward primitive independent calculations, and each core is much weaker than a moden CPU which has a whole bunch of L caches with some prediction behavior systems for code divergence. As you can see on Figure \ref{fg:threadload}, unbalanced thread load assumes for GPU kernel to check every time if the target atom can interact with other atom and if it has to wait till other threads finish their jobs. Thus, we had to come up with a strategy of making equally balanced thread jobs and distribute them among the threads leaving atom to thread mapping as is.
